---
title: "Project 2"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MVA)


```
2) Clean data, detect outliers, and visualizations

First, rather than eliminate rows with missing data, we use value insertion to fill in missing values. This allows us to keep all samples even if they are missing just one data point but still have accurate information for the rest of them.

```{r}
nba <- read.csv("nba.csv")

library(mvnmle)


nba2 <- nba
for(q in 3:ncol(nba2)){
 nba2[is.na(nba2[, q]), q] <- median(nba2[, q], na.rm = TRUE)
}


```

Next, we look at potential outliers for some of our player attributes, and visualize them with a bivariate boxplot

```{r}
library(MVA)
mydata <- nba2[, c("Height", "TRB")]
mydata2 <- nba2[, c("Height", "AST")]
mydata3 <- nba2[, c("FT", "X3P")]
mydata4 <- nba2[, c("PTS", "FG")]
mydata5 <- nba2[, c("AST", "STL")]
bvbox(mydata, xlab = "Height", ylab = "Total Rebounds (per 36 minutes")
bvbox(mydata2, xlab = "Height", ylab = "Assists (per 36 minutes)")
bvbox(mydata3, xlab = "Free Throw Percentage", ylab = "3 Point Percentage (per 36 minutes)")
bvbox(mydata4, xlab = "Points", ylab = "FG Percentage (per 36 minutes)")
bvbox(mydata5, xlab = "Assists", ylab = "Steals")


```

We see some outliers in the rebound, steals, assists and points groups. We remove the extreme outliers in order to make our correlation coefficients more accurate
```{r}
outliers <- match(c("Zhou Qi", "Alan Williams",
                   "Terrence Jones", "Tyler Davis", "Trevon Duval", "Kostas Antetokounmpo", "Gary Payton", "Zach Lofton"),
                 nba2$Player)
nba2 <- nba2[-outliers,]

```

Next, we run a scatterplot matrix to plot all numeric variables and their linear associations, in order to identify relationships. We also add in regression lines for linear fits

```{r}

plot(nba2)
pairs(nba2,
panel = function (x, y, ...) {
points(x, y, ...) 
abline(lm(y ~ x), col = "grey")
 }, pch = ".", cex = 1.5)

```

Next, we run some 3d scatter plots in different groupings to see how the variables relate 


```{r}
library("scatterplot3d")

scatterplot3d(nba2$FT, nba2$X3P, nba2$PTS,
 ,type="h",angle = 40)

scatterplot3d(nba2$STL, nba2$AST, nba2$FG,
 ,type="h",angle = 40)

scatterplot3d(nba2$BLK, nba2$Height, nba2$Age,
 ,type="h",angle = 40)


```

Analyzing the data here, we see that three factors provides useful data with a p value > .05 (where as 4 provides less than .05). Because of the lack of association with a .5 cutoff, we moved the threshold to .25 and gain (expected) results. We seem to be able to eliminate Age as a contributable dimension.

We see the three factors here:

Factor 1: The Big Man - the correlation associates positivesly with taller players having more blocks and rebounds, but not good at free throws and terrible at three pointers.

Factor 2: Sharp Shooter - being the only factor that associates positively enough with the three, we see this leads to a massive correlation with points scored.

Factor 3: Floor General - We see that the shorter players focus on assists and steals.

```{r}
nba.fa <- factanal(nba2[,3:12],factors = 3)
nba.fa$loadings
print(nba.fa$loadings, cutoff = .25)
options(digits = 2)

```

We use those same results and again get the expected outcomes, with a large proportion of those players in Factor 1 being Centers. The other two being split

```{r}
library(mclust)
mc <- Mclust(nba2[,3:12],3)
table(mc$classification, nba2$Pos)
plot(mc, what = "classification")

km <- kmeans(scale(nba2[,3:12]), 3, nstart = 5)
table(km$cluster, nba2$Pos)
plot(nba2[,3:12], col = km$cluster, pch = km$cluster)

```


